---
title: "RdatAI190410"
output: html_notebook
---
Data from Agroinfiltration 190410, measured 5 dpi 190415. Strong hypersensitive response in 9-18, 9-19, 9-20 treatments. Treatments with fewest signs were sampled and measured.  


```{r}
# EC36309data1 <- read.table("190410Summary.csv", header=TRUE,
#                          sep=",") # fluorescence results AI190410 36309 + co-infiltrations
# attach(EC36309data1)
# 
# boxplot(GFP~ID, data=EC36309data1, main="GFP",horizontal=TRUE, varwidth=TRUE, xlab="RFU", ylab="Treatment", ylim=c(0, 45))
# boxplot(RFP~ID, data=EC36309data1, main="RFP", horizontal=TRUE, varwidth=TRUE, xlab="RFU", ylab="Treatment", ylim=c(0, 70))
# boxplot(GFP.RFP~ID, data=EC36309data1, main="GFP to RFP", horizontal=TRUE, varwidth=TRUE, xlab="ratio", ylab="Treatment", ylim=c(0, 2))
# 
# sum(with(EC36309data1, ID=="9"))
# sum(with(EC36309data1, ID=="9.18"))
# sum(with(EC36309data1, ID=="9.19"))
# sum(with(EC36309data1, ID=="9.2"))
# sum(with(EC36309data1, ID=="9.25"))
# sum(with(EC36309data1, ID=="9.28"))
# summary(EC36309data1)

```

Data from 9-18, 9-19, 9-20 and 9-25 should not be trusted, hypersensitive response in most samples most likely skewed RFU data!!

Corrected data: 36309 w/ GFP signal lower than 3.5 RFU were deleted from dataset:
```{r}
# EC36309data2 <- read.table("190410SummaryCorr.csv", header=TRUE,
#                          sep=",") # fluorescence results AI190410 36309 (co-)infiltrations corrected 
# attach(EC36309data2)
# 
# boxplot(GFP~ID, data=EC36309data2, main="GFP",horizontal=TRUE, varwidth=TRUE, xlab="RFU", ylab="Treatment", ylim=c(0, 45))
# boxplot(RFP~ID, data=EC36309data2, main="RFP", horizontal=TRUE, varwidth=TRUE, xlab="RFU", ylab="Treatment", ylim=c(0, 70))
# boxplot(GFP.RFP~ID, data=EC36309data2, main="GFP to RFP", horizontal=TRUE, varwidth=TRUE, xlab="ratio", ylab="Treatment", ylim=c(0, 2))
# 
# sum(with(EC36309data2, ID=="9"))

```
Do not use for data analysis!

-----------------------------------
Data from 7, 7-21, 7-22 co-infiltrations:
```{r}
library(ggplot2)
library(reshape2)

#Combined data from AI190330 (n=6) and AI190410 (n=5)

EC36307_df <- read.table("190410-2.csv", header=TRUE,
                         sep=",")
attach(EC36307_df)

# changing the df to long format

EC36307_df_long <- melt(EC36307_df,
                        id.vars=c("ID"),
                        measure.vars=c("GFP", "RFP", "GFP.RFP" ),
                        variable.name="Channel",
                        value.name="RFU")

attach(EC36307_df_long)

# taking out GFP/RFP ratios
EC36307_df_long_red <- EC36307_df_long[-c(67:99), ] 
EC36307_df_long_red

# taking out GFP, RFP single values
EC36307_df_long_red2 <- EC36307_df_long[-c(1:66), ] 
EC36307_df_long_red2

# taking out treatment 7
EC36307mod_df <- read.table("190410-2mod.csv", header=TRUE,
                         sep=",")
attach(EC36307mod_df)

EC36307mod_df_long <- melt(EC36307mod_df,
                        id.vars=c("ID"),
                        measure.vars=c("GFP", "RFP", "GFP.RFP" ),
                        variable.name="Channel",
                        value.name="RFU")
attach(EC36307mod_df_long)

# taking out GFP/RFP ratios
EC36307mod_df_long_red <- EC36307mod_df_long[-c(45:66), ] 
EC36307mod_df_long_red

```


Boxplots:
```{r}

# GFP vs. RFP
ggplot(data=EC36307_df_long_red, aes(x=ID,y=RFU, colour=factor(Channel))) +
  geom_boxplot(position=position_dodge2(width=0.5), width=0.5) +
  labs(tag = "GFP vs. RFP") +
  scale_color_brewer(palette = "Set2") +
  theme_classic() +
  #facet_wrap(~Channel) +
  #coord_fixed(ratio = 0.0015, xlim = NULL, ylim = NULL) +
  theme(#axis.line = element_line(colour = "black", size = 1, linetype = "solid"),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.text.x= element_text(size=12),
        axis.text.y = element_text(size=12)
        )


#Plot of only 7-21, 7-22
ggplot(data=EC36307mod_df_long_red, aes(x=ID,y=RFU, colour=factor(Channel))) +
  geom_boxplot(position=position_dodge2(width=0.5), width=0.5) +
  labs(tag = "GFP vs. RFP") +
  scale_color_brewer(palette = "Set2") +
  theme_classic() +
  #facet_wrap(~Channel) +
  theme(#axis.line = element_line(colour = "black", size = 1, linetype = "solid"),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.text.x= element_text(size=12),
        axis.text.y = element_text(size=12)
        )


library(reshape2)
library(ggpubr)

my_comparisons <- list( c("T7", "T7-21"), c("T7", "T7-22"))

#GFP/RFP ratios
ggplot(data=EC36307_df_long_red2, aes(x=ID,y=RFU, colour=factor(ID))) +
  geom_boxplot(position=position_dodge(width=0.5), width=0.5) +
  labs(tag = "EC36307") +
  scale_color_grey() + 
  theme_classic() +
  coord_fixed(ratio = 0.18, xlim = NULL, ylim = NULL) +
  #stat_compare_means(method = "t.test", label = "p.format") +
  theme(#axis.line = element_line(colour = "black", size = 1, linetype = "solid"),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.text.x= element_text(size=12),
        axis.text.y = element_text(size=12)
        ) +
  stat_compare_means(method = "wilcox.test", label = "p.format", comparisons = my_comparisons)




```
Paired boxplots:
```{r}
# paired boxplots 
library('dplyr')

t7_gr_df <- filter(EC36307_df_long_red, ID == "T7")
t7_gr_df

t721_gr_df <- filter(EC36307_df_long_red, ID == "T7-21")
t721_gr_df

t722_gr_df <- filter(EC36307_df_long_red, ID == "T7-22")
t722_gr_df

library(ggpubr)

pt7 <-  ggpaired(t7_gr_df, x = "Channel", y = "RFU", color = "Channel", width = 0.5, line.color = "gray", line.size = 0.5) +
  #geom_boxplot(position=position_dodge(width=0.4), width=0.4) +
  labs(tag = "EC36307") +
  scale_colour_brewer(palette = "Set2") +
  theme_classic() +
  #coord_fixed(ratio = 0.25, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme(axis.line = element_line(colour = "black", size = 1, linetype = "solid"),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.x= element_text(size=12),
        axis.text.y = element_text(size=12)
        ) #+
  #stat_compare_means(method = "t.test", label = "p.format", paired = TRUE)

pt721 <- ggpaired(t721_gr_df, x = "Channel", y = "RFU", color = "Channel", width = 0.5, line.color = "gray", line.size = 0.5) +
  #geom_boxplot(position=position_dodge(width=0.4), width=0.4) +
  labs(tag = "EC36307-21") +
  scale_colour_brewer(palette = "Set2") +
  theme_classic() +
  #coord_fixed(ratio = 0.05, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme(axis.line = element_line(colour = "black", size = 1, linetype = "solid"),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.x= element_text(size=12),
        axis.text.y = element_text(size=12)
        )

pt722 <- ggpaired(t722_gr_df, x = "Channel", y = "RFU", color = "Channel", width = 0.5, line.color = "gray", line.size = 0.5) +
  #geom_boxplot(position=position_dodge(width=0.4), width=0.4) +
  labs(tag = "EC36307-22") +
  scale_colour_brewer(palette = "Set2") +
  theme_classic() +
  #coord_fixed(ratio = 0.05, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme(axis.line = element_line(colour = "black", size = 1, linetype = "solid"),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.x= element_text(size=12),
        axis.text.y = element_text(size=12)
        )

pt7
pt721
pt722


```




Statistical testing:
```{r}
############################ Comparison of Means: #############################

## Normality test: qq-plots and Shapiro-Wilk
# Separating all three datasets
EC36307_df_1sam7 <- EC36307_df_long_red2[-c(1:22), ] # 7 
EC36307_df_1sam721 <- EC36307_df_long_red2[-c(12:33), ] # 7-21
EC36307_df_1sam722 <- EC36307_df_long_red2[-c(1:11,23:33), ] #7-22

library(ggpubr)
qq7 <- ggqqplot(data=EC36307_df_1sam7$RFU)
qq7

shapiro.test(EC36307_df_1sam7$RFU)
# 7 does not appear to be normally distributed


qq721 <- ggqqplot(data=EC36307_df_1sam721$RFU)
qq721

shapiro.test(EC36307_df_1sam721$RFU)
# 7-21 does not appear to be normally distributed

qq722 <- ggqqplot(data=EC36307_df_1sam722$RFU)
qq722

shapiro.test(EC36307_df_1sam722$RFU)
# 7-22 possibly normally distributed

## Testing for equal variances:
var(EC36307_df_1sam7$RFU)
var(EC36307_df_1sam721$RFU)
var(EC36307_df_1sam722$RFU)
# reducing df to two sample-df:

EC36307_df_2sam1 <- EC36307_df_long_red2[-c(1:11), ]  # 7, 7-22
EC36307_df_2sam1
EC36307_df_2sam2 <- EC36307_df_long_red2[-c(12:22), ] # 7, 7-21
EC36307_df_2sam2

# F-test for equal variances:
res.ftest1 <- var.test(RFU ~ ID, data = EC36307_df_2sam1)
res.ftest2 <- var.test(RFU ~ ID, data = EC36307_df_2sam2)

res.ftest1
res.ftest2

# -> p-values indicate unequal variances. Normality and equal variances not given. Use non-parametric test.

## Wilcoxon rank-sum test:


wilcox.test(RFU ~ ID, data=EC36307_df_2sam1, alternative = c("two.sided")) 

wilcox.test(RFU ~ ID, data=EC36307_df_2sam2, alternative = c("two.sided")) 

# p-values of 1.134e-05 for both tests: populations are not the same. (There is a minimum p-value in the Wilc. test, possibly that is why both p-values are the same.)
```


Regression:
```{r}
reg7 <- ggplot(EC36307_df, aes(x= RFP, y= GFP, shape = ID, colour=ID)) + #
  labs(x="RFP (RFU)", y="GFP (RFU)") +
  #scale_x_continuous(breaks = pretty(RFP, n = 5)) +
  #scale_y_continuous(breaks = pretty(GFP, n = 5)) +
  geom_point(size=2) + 
  geom_smooth(method=lm, size=0.5) +
  scale_color_brewer(type = 'div', palette = "Accent") +
  #scale_color_grey() +
  theme_classic() +
  coord_fixed(ratio = 0.05, xlim = c(0:50), ylim = c(0:350), expand = TRUE, clip = "on") +
  theme(axis.line = element_line(colour = "black", size = 1, linetype = "solid"),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.text.x= element_text(size=12),
        axis.text.y = element_text(size=12),
        panel.border = element_rect(size = 1, colour = "black", fill = NA)
        ) +
  guides(fill = "none")

reg7

reg8 <- ggplot(EC36307_df, aes(RFP, GFP, shape=ID, colour=ID)) + #
  labs(x="RFP (RFU)", y="GFP (RFU)") +
  #scale_x_continuous(breaks = pretty(RFP, n = 5)) +
  #scale_y_continuous(breaks = pretty(GFP, n = 5)) +
  geom_point( size=2) + 
  geom_smooth(method=lm, size=0.5) +
  #scale_color_brewer(type = 'div', palette = "Spectral") +
  scale_color_grey() +
  theme_classic() +
  guides(fill = "none") +
  #coord_fixed(ratio = 0.05, xlim = c(0:50), ylim = c(0:350), expand = TRUE, clip = "on") +
  theme(axis.line = element_line(colour = "black", size = 1, linetype = "solid"),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.text.x= element_text(size=12),
        axis.text.y = element_text(size=12),
        panel.border = element_rect(size = 1, colour = "black", fill = NA)
        )

reg8

#plotting without extreme values of 7 (- 6 samples):
EC36307_df_low <- EC36307_df[-c(28:33), ] 
EC36307_df_low # n = 5

ggplot(EC36307_df_low, aes(RFP, GFP, shape=ID, colour=ID)) + #
  labs(x="RFP (RFU)", y="GFP (RFU)") +
  #scale_x_continuous(breaks = pretty(RFP, n = 5)) +
  #scale_y_continuous(breaks = pretty(GFP, n = 5)) +
  geom_point( size=2) + 
  geom_smooth(method=lm, size=0.8, se=FALSE, fullrange=FALSE) +
  #scale_color_brewer(
  scale_color_grey() +
  theme_classic() +
  guides(fill = "none") +
  #coord_fixed(ratio = 0.05)+#, xlim = c(0:50), ylim = c(0:350), expand = TRUE, clip = "on") +
  theme(axis.line = element_line(colour = "black", size = 1, linetype = "solid"),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.text.x= element_text(size=12),
        axis.text.y = element_text(size=12),
        panel.border = element_rect(size = 1, colour = "black", fill = NA)
        )

```

SLOPE COMPARISON
```{r}
##############################################################
#           SLOPE COMPARISON
##############################################################
# Scatterplot of GFP to RFP for 7 n = 5, (lowest 5 datapoints), 7-21 and 7-22
library(devtools)
source_gist("524eade46135f6348140")
ggplot(EC36307_df_low, aes(RFP, GFP, shape=ID, colour=ID)) + #
  labs(x="RFP (RFU)", y="GFP (RFU)") +
  #scale_x_continuous(breaks = pretty(RFP, n = 5)) +
  #scale_y_continuous(breaks = pretty(GFP, n = 5)) +
  geom_point( size=2) +
  stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE) + #show equations in figure
  geom_smooth(method=lm, size=0.8, se=TRUE, alpha = 0.15, fullrange=FALSE) +
  #scale_color_brewer(
  scale_color_grey() +
  theme_classic() +
  guides(fill = "none") +
  #coord_fixed(ratio = 0.05)+#, xlim = c(0:50), ylim = c(0:350), expand = TRUE, clip = "on") +
  theme(axis.line = element_line(colour = "black", size = 1, linetype = "solid"),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.text.x= element_text(size=12),
        axis.text.y = element_text(size=12),
        panel.border = element_rect(size = 1, colour = "black", fill = NA)
        )


# Comparing slopes, EC36307 only n = 5 (lowest five values) 
library(lsmeans)
m.interaction <- lm(GFP ~ RFP*ID, data = EC36307_df_low)
m.interaction
# ANOVA not possible due to non-normality and heteroscedascdicity


# Obtain slopes
m.interaction$coefficients
m.lst <- lstrends(m.interaction, "ID", var="RFP")
m.lst

# Compare slopes
pairs(m.lst)
##############################################################

```

```{r}
####### SLOPE COMPARISON: All samples included
m.interaction2 <- lm(GFP ~ RFP*ID, data = EC36307_df) # prediction of GFP (dependent) signal through RFP (independent) according to ID

# Obtain slopes
m.interaction2$coefficients
m.lst2 <- lstrends(m.interaction2, "ID", var="RFP")
m.lst2

# Compare slopes
pairs(m.lst2)

# high p-values: including large datapoints of 7 made all interactions non-significant.

#visualize by scatterplot
ggplot(EC36307_df, aes(RFP, GFP, shape=ID, colour=ID)) + #
  labs(x="RFP (RFU)", y="GFP (RFU)") +
  #scale_x_continuous(breaks = pretty(RFP, n = 5)) +
  #scale_y_continuous(breaks = pretty(GFP, n = 5)) +
  geom_point( size=2) +
  stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE) + #show equations in figure
  geom_smooth(method=lm, size=0.8, se=TRUE, alpha = 0.15, fullrange=FALSE) +
  #scale_color_brewer(
  scale_color_grey() +
  theme_classic() +
  guides(fill = "none") +
  coord_fixed(ratio = 0.1, xlim = c(0:50), ylim = c(0:300), expand = TRUE, clip = "on") +
  theme(axis.line = element_line(colour = "black", size = 1, linetype = "solid"),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.text.x= element_text(size=12),
        axis.text.y = element_text(size=12),
        panel.border = element_rect(size = 1, colour = "black", fill = NA)
        )


```



Regression summary: p-values:
```{r}
summary(m.interaction)
summary(m.lst)
```

