---
title: "R Data Agroinfiltration 190322"
output: html_notebook
pdf_document: Rnotebook_190322
---

Date: 190327
Sampled: 15 plants 5 days post infection (dpi), one leaf disc per plant for treatment w/ EC36309, coinfiltration treatments 9-18, 9-19, 9-20 (EC36309 w/ EC36318/19/20). Co-infiltration data unusable due to strong hypersensitive response in leaf spots.

Two separate technical replicates of 36309 endpoint fluorescence measurement, data from AI190320 and AI190322.

Analysis of EC36309 using ggplot2:
```{r}
merged9_df <- read.table("mergedEC36309.csv", header=TRUE,
                         sep=",")

#merged dataset of AI190320 (n = 15) and AI190322 (n=15), only analyzing EC36309 treatment (no observable HR in samples).

attach(merged9_df)
summary(merged9_df)

library(ggplot2)
library(reshape2)

# changing the df to long format

merged9_df_long <- melt(merged9_df,
                        id.vars=c("ID"),
                        measure.vars=c("GFP", "RFP", "GFP.RFP" ),
                        variable.name="Channel",
                        value.name="RFU")

attach(merged9_df_long)

# taking out GFP/RFP ratios for GFP to RFP comparisons

merged9_df_long_red <- merged9_df_long[-c(61:90), ] 
merged9_df_long_red

ggplot(data=merged9_df_long_red, aes(x=Channel,y=RFU)) +
  geom_boxplot(position=position_dodge(width=0.4), width=0.4) +
  labs(tag = "EC36309") +
  scale_color_grey() + 
  theme_classic() +
  coord_fixed(ratio = 0.05, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme(axis.line = element_line(colour = "black", size = 1, linetype = "solid"),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.x= element_text(size=12),
        axis.text.y = element_text(size=12)
        )

summary(merged9_df_long_red)

# df with only GFP/RFP ratios:

merged9_df_long_red2 <- merged9_df_long[-c(1:60), ] 
merged9_df_long_red2

ggplot(data=merged9_df_long_red2, aes(x=Channel,y=RFU)) +
  geom_boxplot(position=position_dodge(width=0.4), width=0.4) +
  #scale_y_continuous(breaks = 1) +
  scale_color_grey() + 
  theme_classic() +
  coord_fixed(ratio = 3, xlim = NULL, ylim = c(0,1), expand = TRUE, clip = "on") +
  theme(axis.line = element_line(colour = "black", size = 1, linetype = "solid"),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.x= element_text(size=12),
        axis.text.y = element_text(size=12)
        )

# Testing normality, qq-plot
library(ggpubr)
qq9 <- ggqqplot(data=merged9_df_long_red2$RFU)
qq9

shapiro.test(merged9_df_long_red2$RFU)

# p-value = 0.4, sample distribution is likely to be normal.
#The null hypothesis of these tests is that "sample distribution is normal". If the test is significant, the distribution is non-normal.

```

Quite stark differences in aboslut values could be normalized by the GFP/RFP ratio between both replicates.
Shapiro-Wilk normality test returns p-value of 0.39, large probability that the dataset is normally distributed. qqplot further underscores this.

Paired box plot:
```{r}
library(ggpubr)
ggpaired(merged9_df_long_red, x = "Channel", y = "RFU", color = "Channel", width = 0.5, line.color = "gray", line.size = 0.5) +
  #geom_boxplot(position=position_dodge(width=0.4), width=0.4) +
  labs(tag = "EC36309") +
  scale_colour_brewer(palette = "Set2") +
  theme_classic() +
  coord_fixed(ratio = 0.05, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme(axis.line = element_line(colour = "black", size = 1, linetype = "solid"),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.x= element_text(size=12),
        axis.text.y = element_text(size=12)
        ) +
  stat_compare_means(method = "t.test", label = "p.format", paired = TRUE)


```


Linear regression in ggplot2:
```{r}
library(ggplot2)
reg2 <- ggplot(merged9_df, aes(x=RFP, y=GFP)) +
  labs(x="RFP (RFU)", y="GFP (RFU)") +
  scale_x_continuous(breaks = pretty(RFP, n = 5)) +
  scale_y_continuous(breaks = pretty(GFP, n = 5)) +
  geom_point(shape=1, size=2) + 
  geom_smooth(method=lm, colour="black", size=0.5) +
  scale_color_grey() +
  theme_classic() +
  coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme(axis.line = element_line(colour = "black", size = 1, linetype = "solid"),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.x= element_text(size=12),
        axis.text.y = element_text(size=12),
        panel.border = element_rect(size = 1, colour = "black", fill = NA)
        )
  #theme(panel.background = element_rect(fill = "white", colour = "black"))

reg2

reg3 <- ggplot(merged9_df, aes(x=RFP, y=GFP.RFP)) +
  labs(x="RFP (RFU)", y="GFP.RFP") +
  scale_x_continuous(breaks = pretty(RFP, n = 5)) +
  scale_y_continuous(breaks = pretty(GFP.RFP, n = 5)) +
  geom_point(shape=1, size=2) + 
  geom_smooth(method=lm, colour="black", size=0.5) +
  scale_color_grey() +
  theme_classic() +
  #coord_fixed(ratio = 30, xlim = c(0:65), ylim = c(0:1)) +
  theme(axis.line = element_line(colour = "black", size = 1, linetype = "solid"),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.x= element_text(size=12),
        axis.text.y = element_text(size=12),
        panel.border = element_rect(size = 1, colour = "black", fill = NA)
        )
  #theme(panel.background = element_rect(fill = "white", colour = "black"))

reg3


reg4 <- ggplot(merged9_df, aes(x=GFP, y=GFP.RFP), ylim = c(0:1)) +
  labs(x="GFP (RFU)", y="GFP.RFP") +
  scale_x_continuous(breaks = pretty(RFP, n = 5)) +
  scale_y_continuous(breaks = pretty(GFP.RFP, n = 5)) +
  geom_point(shape=1, size=2) + 
  geom_smooth(method=lm, colour="black", size=0.5) +
  scale_color_grey() +
  theme_classic() +
  #coord_fixed(ratio = 15, xlim = c(1:25), ylim = c(0:1), expand = TRUE, clip = "on") +
  theme(axis.line = element_line(colour = "black", size = 1, linetype = "solid"),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.x= element_text(size=12),
        axis.text.y = element_text(size=12),
        panel.border = element_rect(size = 1, colour = "black", fill = NA)
        )
  #theme(panel.background = element_rect(fill = "white", colour = "black"))

reg4





```

Correlating GFP to RFP in reference measurement EC36309, n = 30. GFP and RFP signals correlate very well. Reporter system seems usable for our hypothesis testing. 
```{r}
scatter.plot.1 <- plot(EC36309merg$GFP, EC36309merg$RFP, main="Scatterplot", xlab="GFP", ylab = "RFP", xlim = c(0,30), ylim = c(0,70))
reg1 <- lm(GFP~RFP,data = merged9_df)
summary(reg1)

with(merged9_df,plot(RFP, GFP))
	abline(reg1)
	
	
	reg2 <- lm(GFP.RFP~RFP, data = merged9_df)
summary(reg2)

reg3 <- lm(GFP.RFP~GFP, data = merged9_df)
summary(reg3)
	
# newx = seq(min(RFP),max(RFP), by = 0.05) #,by = 1.12
# conf_interval <- predict(reg1, newdata=data.frame(x=newx), interval="confidence",
#                          level = 0.95)
# plot(RFP, GFP, xlab="RFP", ylab="GFP", main="Regression")
# abline(reg1, col="black")
# matlines(newx, conf_interval[,2:3], lty=2, col = "blue")
```


Power analysis:
```{r}
#install.packages("pwr")
library(pwr)
pwr.t.test(n = 30, d = NULL, sig.level = 0.05, power = 0.8, 
           type = "two.sample",
           alternative = "greater")

pwr.t.test(n = 15, d = NULL, sig.level = 0.01, power = 0.99, 
           type = "two.sample",
           alternative = "greater")

```
Power calculation for a two-sample t-test, sample size of 30, alpha of 5%, power of 80% gives an effect size of 0.65.
Power calculation for a two-sample t-test, sample size of 15, alpha of 1%, power of 99% gives an effect size of 1.72. 

```{r}
VarGFP.RFP <- var(EC36309merg$GFP.RFP)
VarGFP.RFP
sdGFP.RFP <- sqrt(VarGFP.RFP)
sdGFP.RFP               #sample standard deviation of GFP/RFP ratio EC36309 = 0.13
sdGFP.RFPlarge <- 0.25  #setting standard deviation to a conservative estimate (almost double from initial)


avgGFPRFPeff <- median(EC36309merg$GFP.RFP*1.50) #detecting x% greater signal of GFP
avgGFPRFP <- median(EC36309merg$GFP.RFP)            #reference

deff <- (avgGFPRFPeff-avgGFPRFP)/sdGFP.RFP          #calculating effect size
deff

dlarge <- (avgGFPRFPeff-avgGFPRFP)/sdGFP.RFPlarge       #calculating effect size for larger sd, assuming we underestimate sd.
dlarge

pwr.conservative <- pwr.t.test(n = NULL, d = dlarge, sig.level = 0.05, power = 0.80, 
           type = "two.sample",
           alternative = "greater") #power calculation under the assumption of larger uncertainty (almost double of estimation)
pwr.conservative
plot(pwr.conservative)

pwr.est <- pwr.t.test(n = NULL, d = deff, sig.level = 0.05, power = 0.8, 
           type = "two.sample",
           alternative = "greater") #power calculation under the assumption of estimated uncertainty
pwr.est
plot(pwr.est)

# What's the difference in ratio that can be caught w/ large n, conservative sd
#GFP.RFPrat <- dlarge * sdGFP.RFPlarge
#GFP.RFPrat 

# Two-tailed t-test instead of one-tailed
pwr.conservative2 <- pwr.t.test(n = NULL, d = dlarge, sig.level = 0.05, power = 0.80, 
           type = "two.sample") #two-tailed power calculation under the assumption of larger uncertainty (almost double of estimation)
pwr.conservative2
plot(pwr.conservative2)

pwr.est2 <- pwr.t.test(n = NULL, d = deff, sig.level = 0.05, power = 0.8, 
           type = "two.sample") #two-tailed power calculation under the assumption of estimated uncertainty
pwr.est2
plot(pwr.est2)

# Power calc. of conservative estimate of sd of 0.2 instead of 0.13
d3 <- (avgGFPRFPeff-avgGFPRFP)/0.2
d3

pwr.conservative3 <- pwr.t.test(n = 10, d = d3, sig.level = 0.05, power = NULL, 
           type = "two.sample") #two-tailed power calculation under the assumption of larger uncertainty (almost double of estimation)
pwr.conservative3


```




ANOVA power test?
```{r}

#example
groupmeans <- c(550, 598, 598, 646)

p <- power.anova.test(groups = 4, 
between.var = var(groupmeans), within.var = 6400, 
power=0.823,sig.level=0.05,n=NULL)
p



```



CRISPR validation:
Treatment: EC36317.2, 4 plants (n = 4), three spots per leaf, one per treatment. No RFP signal, therefore data inconclusive.
```{r}
EC36317dat1 <- read.table("AI190322EC36317.csv", header=TRUE,
                         sep=",") #data from 36317 treatnent from AI190322 
boxplot(GFP~ID, data=EC36317dat1, main="GFP", horizontal=TRUE, varwidth=TRUE, xlab="RFU", ylab="Treatment", ylim=c(0, 200))
boxplot(RFP~ID, data=EC36317dat1, main="RFP", horizontal=TRUE, varwidth=TRUE, xlab="RFU", ylab="Treatment", ylim=c(0, 45))
```

EC36317.2 showed again only GFP expression and no RFP signal, the construct and glycerol stocks were discarded. Co-infiltration with CRISPR constructs EC36321 and 36322 targeting mNeonGreen and tdTomato CDS respectively. 17-21 showed no GFP expression, treatment with construct targeting mNeonGreen. Without tdTomato reference, this could still be just an artifact from improper overall expression of the construct. 








